Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:
helpful when exploring two continuous variables


```{r Scatterplots}

#setwd("W:/daniel.scott/Udacity/Lesson5_TwoVariables")
install.packages("ggplot2")
library(ggplot2)
# can also use  this to read in data
#read.delim('pseudo_facebook.tsv')
pf <- read.csv("pseudo_facebook.tsv", sep ='\t')

qplot(x = age, y=friend_count, data=pf)
```

***

#### What are some things that you notice right away?
Response:
left skewed data. People lower than 30 have on average highest friend count. Then there are sporadically high friend counts for people approx. 65 and 100

***

### ggplot Syntax
Notes:
differs from qplot as it can deal with more complexity. 
add the x and y inside the aes()
reccomended to add a layer one at a time to allow for debugging
can add a limit to the axis as additional layer
```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data =pf) + geom_point()

summary(pf$age)
#minimum is 13 max is 113 but there are doubts thats real so lets put 90 as the max
ggplot(aes(x = age, y = friend_count), data =pf) + geom_point() + xlim(13,90)
```

***

### Overplotting
Notes:
Overplotting is when its difficult to be able to tell how many points are in each region
Overcome this with alpha parameter in ggplot. in the below we state that it will take 20 points to colour the same as 1 in the plot above 

geom_jitter can replace geom_pointand spread the values over 
```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data =pf) + geom_point(alpha = 1/20) + xlim(13,90)

ggplot(aes(x = age, y = friend_count), data =pf) + geom_jitter(alpha = 1/20) + xlim(13,90)
```

#### What do you notice in the plot?
Response:
The majority of the data falls below the 1000 friend count
two peaks around 20 and then 69 
***

### Coord_trans()
Notes:
https://ggplot2.tidyverse.org/reference/coord_trans.html#arguments
```{r Coord_trans()}

ggplot(aes(x=age, y=friend_count), data = pf) + geom_jitter(alpha = 1/20) + xlim(13,90)
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

Beware when sqrt on 0 values as we could possibly get negative numbers.
Can add a position value to the geom_point by adding a height of 0

```{r}
ggplot(aes(x=age, y=friend_count), data = pf) + geom_point(alpha = 1/20, position = position_jitter(h = 0)) + xlim(13,90)  + coord_trans(y="sqrt")
```

#### What do you notice?
Much easier to see friend count and age
***

### Alpha and Jitter
Notes:
explore relationship of friends initiated vs age
```{r Alpha and Jitter}
qplot(x = age, y=friendships_initiated, data=pf)
summary(pf$friendships_initiated)

ggplot(aes( x = age, y = friendships_initiated), data = pf) + geom_point(alpha = 1/10, position = position_jitter(h = 0)) + xlim(13,90)  + coord_trans(y="sqrt")
```

***

### Overplotting and Domain Knowledge
Notes:
sometimes it makes more sense to think about % rather than absolute values especially when there are large differences that can give poor interpretation 

***

### Conditional Means
Notes:
its good to see the summary of two variables to ask questions i.e. how does friend count vary with age?
dpylr lets us split up dataframes and apply function to that section
```{r Conditional Means}
install.packages('dplyr')
library(dplyr)
#filter()
#groupby()
#mutate()
#arrange()
#we can group the data by the age
age_groups <- group_by(pf, age)
#create variable which we will summarise, here we want the summary of age_groups
pf.fc_by_age <- summarise(age_groups, 
                          #find out the mean
                          friend_count_mean = mean(friend_count),
                          #find out the median
                          friend_count_median = median(friend_count),
                          #n function only used in summarise, counts numbers in each group
                          n = n())

head(pf.fc_by_age)

#dataframe isnt in order so we use arrange to arrange the data from low to high

pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age)


###alternative way
# %>% chain functions onto datasets, perform one function at a time one after the other on pf

pf.fc_by_age2 <- pf%>%
  group_by(age)%>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n())%>%
  arrange(age)

head(pf.fc_by_age2)

```

Create your plot!
plot av friend count vs age
```{r Conditional Means Plot}
#ggplot(aes( x = age, y = friend_count_mean), data = pf.fc_by_age2) #+ geom_point(alpha = 1/10, position = position_jitter(h = 0)) + xlim(13,90)  + coord_trans(y="sqrt")
library(ggplot2)
ggplot(data = pf.fc_by_age2,aes(x = age, y =friend_count_median)) + geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes( x = age, y = friendships_initiated), data = pf) + geom_point(alpha = 1/10, position = position_jitter(h = 0), color = 'orange') + xlim(13,90) +  coord_trans(y="sqrt") +
geom_line(stat = "summary", fun.y = mean) +
  #get the 10% quartile use 0.1 in probs and quartile function note the use of fun.args
geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = 0.1), linetype =2, color = 'blue') +
  #90% quantile
geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = 0.9), linetype =2, color = 'blue') +
# add the median line in the fun.y 
stat_summary(fun.y = median, geom="line", color = "purple")
  #can add coord_cartesian layers but need to remove the xlim and coord_trans layer in line 167
  #coord_cartesian(xlim = c(60, 80))
```

#### What are some of your observations of the plot?
Response:
having more than 1000 friends is rare

90th percentile is well below 1000

35 - 60 year olds have 90th percentile lower than 250

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:
created histograms to show summary data by creating histogram charts showing how people over/underestimated audience. Clearly see how the majority of facebook users underestimate their audience 
***

### Correlation
Notes:
http://www.statisticssolutions.com/correlation-pearson-kendall-spearman/
Rule of thumb, correlation > 0.3 | -0.3< is meaningful, 0.5 is moderate and 0.7 is large
"Pearson" method measures the relationship between two different variables
```{r Correlation}
with(pf, cor.test(age, friend_count))

#or
cor.test(pf$age, pf$friend_count, method = "pearson")
"""Pearson's product-moment correlation

data:  pf$age and pf$friend_count
t = -8.6268, df = 99001, p-value < 2.2e-16
alternative hypothesis: true correlation is not equal to 0
95 percent confidence interval:
 -0.03363072 -0.02118189
sample estimates:
        cor -0.02740737"""
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response: -0.02740737

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age,friend_count))

```

***

### Correlation Methods
Notes:
monotonic relationships use spearmen correlation
https://en.wikipedia.org/wiki/Monotonic_function
Spearman rank correlation is a non-parametric test that is used to measure the degree of association between two variables
Kendall rank correlation is a non-parametric test that measures the strength of dependence between two variables.  
***
```{r}
with(subset(pf, age <= 70), cor.test(age,friend_count, method = "spearmen"))

```

## Create Scatterplots
Notes:
There are some extreme outliers in this data set
```{r}
summary(pf$likes_received)

summary(pf$www_likes_received)

quantile(pf$www_likes_received, 0.90)

quantile(pf$likes_received, 0.90)

with(subset(pf, likes_received < 100), cor.test(likes_received,www_likes_received))

ggplot(aes(x=likes_received, y=www_likes_received), data = pf) + geom_point() + xlim(0,300) +ylim(0,100)
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

