Lesson 3
========================================================
qplot cheatsheet
https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf

Reviewing outliers, missing values, distributions, boxplots, histograms, freq polygons
Limits, bin widths
data in  log and binary 
***

### What to Do First?
Notes:

getwd()
list.files()
***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
setwd('C:/Users/daniel.scott/downloads')
#https://s3.amazonaws.com/udacity-hosted-downloads/ud651/pseudo_facebook.tsv
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
#names provides the output of variables (columns)
names(pf)
```
#

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
library(ggplot2)
#adding label for each bin by adding the scale breaks 1 to 31
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks = 1:31)
```

***

#### What are some things that you notice about this histogram?
Response:
Unusually large bin for the 1st. 
***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes: You want to do split up your data by one or more variables and plot the subsets of data together.

It splits up the data according to a different variable i.e. below we can see the dataset which is shows the freq of birthdays and with a facet we can see th freq of days per month 

http://www.cookbook-r.com/Graphs/Facets_(ggplot2)/
```{r Faceting}
qplot(x = dob_day, data = pf) +
     scale_x_continuous(breaks = 1:31) +
     facet_wrap(~dob_month, ncol = 3)
```

#### Let's take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira's outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x = friend_count, data = pf)


```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:
xlim limits the x max to plot
or scale_x_con with limits does the same

```{r Limiting the Axes}

qplot(x = friend_count, data = pf, xlim = c(0,1000))

```

### Exploring with Bin Width
Notes:
Skewed data to the left side of the graph

***

### Adjusting the Bin Width
Notes:
https://ggplot2.tidyverse.org/reference/scale_continuous.html

binwidth can limit the bin size
can also add facegrid or facet wrap
### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
#can use facet grid too
#  facet_grid(gender ~ .)
```

***

### Omitting NA Values
Notes:
subset data to remove NA - using !is.na(gender) means if in gender there are na's dont include them
if you use na.omit it removes any row that contains NA, not just for gender
```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(lim = c(0,1000), breaks = seq(0,1000,50)) + 
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:
by using table(pf$gender) you can see the total number of inputs for each, 

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:
females
#### What's the difference between the median friend count for women and men?
Response:
Female has 22 more median friends

#### Why would the median be a better measure than the mean?
Response:
because the mean includes outliers which can skew the data. median is resistant to change as it marks the halfway point for all data points

***

### Tenure
Notes:

```{r Tenure}
qplot(x = tenure, data = pf, color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(lim = c(0,2000), breaks = seq(0,2000,200))

```

***

#### How would you create a histogram of tenure by year?
as tenure is in days we can convert it to years by dividing by 365
```{r Tenure Histogram by Year}
qplot(x = tenure/365, data = pf, binwidth = 1) +
  scale_x_continuous(lim = c(0,10), breaks = seq(0,10,1))
```

***

### Labeling Plots
Notes:
R uses the variables as default labels so you can add labels by using xlab and ylab

```{r Labeling Plots}
qplot(x = tenure/365, data = pf,
      xlab = '# years using Facebook',
      ylab = '# users in sample',
      color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(lim = c(0,7), breaks = seq(1,7,1))
```

***

### User Ages
Notes:
Can find out the maximum age by using summary(pf), binwidth =1 because its age data
```{r User Ages}
qplot(x = age, data = pf, binwidth = 1,
      ylab = '# users in sample',
      color = I('black'), fill = I('#F79420'))
      scale_x_continuous(breaks = seq(0,113,5))
```

#### What do you notice?
Response:
Untrustable data over 90 years old
***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:
summary(log10(pf$friend_count +1))
gives the log 10 of the friend count and to avoid -infinity minimum we add 1 
***

### Add a Scaling Layer
Notes:
http://lightonphiri.org/blog/ggplot2-multiple-plots-in-one-graph-using-gridextra
install.packages("gridExtra") #install grid arrange package
define the basic graph then create other by adding new layers
library(gridExtra)
main= adds title 
if you have a long tail graph its alwas helpful to use log scaling to be able to see details in the tail
```{r Add a Scaling Layer}
#with titles on the graph use below code
p1 <- qplot( x=friend_count, data = pf, main = "facebook data")
p2 <- qplot(x = log10(friend_count +1), data = pf)
p3 <- qplot(x = sqrt(friend_count), data = pf)

library(gridExtra)
#using ncol you can arrange the graphs to be in one column or more
grid.arrange(p1, p2, p3, ncol =1)

#can also use ggplot
#aes means aesthetic, geom is the type of graph
p4 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p5 <- p4 + scale_x_log10()
p6 <- p4 + scale_x_sqrt()


#important to note the difference in scale labelling with 

#qplot gives the log value
p2 <- qplot(x = log10(friend_count +1), data = pf)

#ggplot gives the actual values 
p5 <- p4 + scale_x_log10()

#can plot the actual values in qplot by using the original p1 + log layer
qplotlog <- p1 + scale_x_log()
```

***


### Frequency Polygons
we can use the same code as when we used facet_wrap but we add the geom to 'freqpoly' and set the colour to differentiate the genders
```{r Frequency Polygons}
library(ggplot2)
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10, geom = 'freqpoly', color = gender) +
    scale_x_continuous(lim = c(0,1000), breaks = seq(0,1000,50))
    
##the y= part calculates the proportion
qplot(x = friend_count, y = ..count../sum(..count..), data = subset(pf, !is.na(gender)), xlab = 'Friend count', ylab = 'Proportion of users with that friend count', binwidth = 10, geom = 'freqpoly', color = gender) +
    scale_x_continuous(lim = c(0,1000), breaks = seq(0,1000,50))
    
    
# from the code below you can zoom into the tail of the graph and see that when fc is over 800 women have a higher proportion than men in friend count and overtake 
qplot(x = friend_count, y = ..count../sum(..count..), 
      data = subset(pf, !is.na(gender)), 
      xlab = 'Friend count', 
      ylab = 'Proportion of users with that friend count', 
      binwidth = 10, geom = 'freqpoly', color = gender) +
     scale_x_continuous(lim = c(400,1000), breaks = seq(0,1000,50))
```

***

### Likes on the Web
Notes:
below we see that women overtake men on likes  
```{r Likes on the Web}
qplot(x = likes_received, data = subset(pf, !is.na(gender)), geom = 'freqpoly', color = gender) +
     scale_x_continuous() +
     scale_x_log10()

#code below shows the overtaking happens around 29 votes
qplot(x = www_likes, data = subset(pf, !is.na(gender)), geom = 'freqpoly', color = gender) +
     scale_x_continuous() +
     scale_x_log10(lim = c(1,100),breaks = seq(0,1000,10))

#if we want to see the total number we can use aggregate which totals up all of the  
aggregate(www_likes ~ gender, data = pf, sum)

#alternatively we use by
by(pf$www_likes, pf$gender, sum)

```
Implications of above - women give nearly 3 million more likes than men, over double - of interest to marketing teams

***

### Box Plots
Notes:

```{r Box Plots}
qplot(x = gender, y = friend_count, data = subset(pf, !is.na(gender)), geom = 'boxplot')
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.
ylim can be used to limit to axis. it removes data from visualisation
coord_cartesian()

```{r}
qplot(x = gender, y = friend_count, data = subset(pf, !is.na(gender)), geom = 'boxplot') +
#ylim(0,1000)
coord_cartesian(ylim =c(0,1000))
```

***

### Box Plots, Quartiles, and Friendships
Notes:
females have more friends that men (line through boxplot is mean) by looking at the summary we can see that the mean for females in 242 and 165 for men

quartiles - can see from the boxplot that females have 3rd quartile of 244 (75% femals have less than 244 friends or 25% females have more than 244 friends) 

```{r Box Plots, Quartiles, and Friendships}
qplot(x = gender, y = friend_count, data = subset(pf, !is.na(gender)), geom = 'boxplot') +
#ylim(0,1000)
  coord_cartesian(ylim =c(0,250))
                
by(pf$friend_count, pf$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:females
```{r Friend Requests by Gender}
qplot(x = gender, y = friendships_initiated, data = subset(pf, !is.na(gender)), geom = 'boxplot') +
+     #ylim(0,1000)
+     coord_cartesian(ylim =c(0,200))
by(pf$friendships_initiated, pf$gender, summary)
```

Response:
75% of males initated less than the mean average of females initiated, with the average of initiating over 10 more friends
***

### Getting Logical
Notes:
if you have alot of zeros in your results, it can be easier to convert your data into binary values (1 for true, 0 for false)
```{r Getting Logical}
#checks how many people have checked in
summary(pf$mobile_likes > 0)

#create a new variable, assigned NA values to the variabel
pf$mobile_check_in <- NA

#change these NAs to 1 if the user has checked in, if they have it assigns 1 otherwise assigns 0
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)

#make the binary values as factor variables
pf$mobile_check_in <- factor(pf$mobile_check_in)

#get the summary of these new values
summary(pf$mobile_check_in)

#to get the pct of users that have checked in use to return the %
length(which(pf$mobile_check_in == 1))/length(pf$mobile_check_in)
```

Response:
As over half of the users have used mobile checkins it shows that it is an important experience for the users and therefore in the companies interest to keep investing in this
***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!
