Lesson 3
========================================================
qplot cheatsheet
https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf
***

### What to Do First?
Notes:

getwd()
list.files()
***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
#https://s3.amazonaws.com/udacity-hosted-downloads/ud651/pseudo_facebook.tsv
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
#names provides the output of variables (columns)
names(pf)
```
#

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
library(ggplot2)
#adding label for each bin by adding the scale breaks 1 to 31
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks = 1:31)
```

***

#### What are some things that you notice about this histogram?
Response:
Unusually large bin for the 1st. 
***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes: You want to do split up your data by one or more variables and plot the subsets of data together.

It splits up the data according to a different variable i.e. below we can see the dataset which is shows the freq of birthdays and with a facet we can see th freq of days per month 

http://www.cookbook-r.com/Graphs/Facets_(ggplot2)/
```{r Faceting}
qplot(x = dob_day, data = pf) +
     scale_x_continuous(breaks = 1:31) +
     facet_wrap(~dob_month, ncol = 3)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x = friend_count, data = pf)


```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:
xlim limits the x max to plot
or scale_x_con with limits does the same

```{r Limiting the Axes}

qplot(x = friend_count, data = pf, xlim = c(0,1000))

```

### Exploring with Bin Width
Notes:
Skewed data to the left side of the graph

***

### Adjusting the Bin Width
Notes:
https://ggplot2.tidyverse.org/reference/scale_continuous.html

binwidth can limit the bin size
can also add facegrid or facet wrap
### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
#can use facet grid too
#  facet_grid(gender ~ .)
```

***

### Omitting NA Values
Notes:
subset data to remove NA - using !is.na(gender) means if in gender there are na's dont include them
if you use na.omit it removes any row that contains NA, not just for gender
```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(lim = c(0,1000), breaks = seq(0,1000,50)) + 
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:
by using table(pf$gender) you can see the total number of inputs for each, 

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:
females
#### What's the difference between the median friend count for women and men?
Response:
Female has 22 more median friends

#### Why would the median be a better measure than the mean?
Response:
because the mean includes outliers which can skew the data. median is resistant to change as it marks the halfway point for all data points

***

### Tenure
Notes:

```{r Tenure}
qplot(x = tenure, data = pf, color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(lim = c(0,2000), breaks = seq(0,2000,200))

```

***

#### How would you create a histogram of tenure by year?
as tenure is in days we can convert it to years by dividing by 365
```{r Tenure Histogram by Year}
qplot(x = tenure/365, data = pf, binwidth = 1) +
  scale_x_continuous(lim = c(0,10), breaks = seq(0,10,1))
```

***

### Labeling Plots
Notes:
R uses the variables as default labels so you can add labels by using xlab and ylab

```{r Labeling Plots}
qplot(x = tenure/365, data = pf,
      xlab = '# years using Facebook',
      ylab = '# users in sample',
      color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(lim = c(0,7), breaks = seq(1,7,1))
```

***

### User Ages
Notes:
Can find out the maximum age by using summary(pf), binwidth =1 because its age data
```{r User Ages}
qplot(x = age, data = pf, binwidth = 1,
      ylab = '# users in sample',
      color = I('black'), fill = I('#F79420'))
      scale_x_continuous(breaks = seq(0,113,5))
```

#### What do you notice?
Response:
Untrustable data over 90 years old
***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:
summary(log10(pf$friend_count +1))
gives the log 10 of the friend count and to avoid -infinity minimum we add 1 
***

### Add a Scaling Layer
Notes:
http://lightonphiri.org/blog/ggplot2-multiple-plots-in-one-graph-using-gridextra
install.packages("gridExtra") #install grid arrange package
define the basic graph then create other by adding new layers
library(gridExtra)
main= adds title 
```{r Add a Scaling Layer}
#with titles on the graph use below code
p1 <- qplot( x=friend_count, data = pf, main = "facebook data")
p2 <- qplot(x = log10(friend_count +1), data = pf)
p3 <- qplot(x = sqrt(friend_count), data = pf)

#using ncol you can arrange the graphs to be in one column or more
grid.arrange(p1, p2, p3, ncol =1)

#can also use ggplot
#aes means aesthetic, geom is the type of graph
p4 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p5 <- p4 + scale_x_log10()
p6 <- p4 + scale_x_sqrt()


#important to note the difference in scale labelling with 

#qplot gives the log value
p2 <- qplot(x = log10(friend_count +1), data = pf)

#ggplot gives the actual values 
p5 <- p4 + scale_x_log10()

#can plot the actual values in qplot by using the original p1 + log layer
qplotlog <- p1 + scale_x_log()
```

***


### Frequency Polygons

```{r Frequency Polygons}

```

***

### Likes on the Web
Notes:

```{r Likes on the Web}

```


***

### Box Plots
Notes:

```{r Box Plots}

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!
